\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-04-18 archive.tex
almost done except for the conclusion!

\subsection{Programs that manipulate sequences}\label{sequences}
We first apply \system to two classic benchmark domains: list
processing and text editing. In both cases we solve tasks specified by
a input/output examples, starting with a generic functional programming
basis, including routines like \code{map}, \code{fold}, \code{cons}, \code{car}, \code{cdr}, etc.
%% : \code{foldr}, \code{unfold}, \code{if}, \code{map},
%% \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0},
%% \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and
%% \code{is-nil}.

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system typically composes around 20 new library routines, and discovered multiple
higher-order functions. Each round of memory consolidation built on
components discovered in earlier sleep cycles --- for example the
agent first learns the higher-order function \code{filter}, uses
\code{filter} to learn to take the maximum element of a list, then
uses that routine to learn a new component for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
solve a task involving sorting a list of numbers (Figure~\ref{exampleDSL}).
This incremental, modular learning of deep hierarchies of library
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the abstraction phase
of sleep).

@@

How do \systemEnding's representations of these problems themselves change, as the model acquires expertise? Within cognitive science, a rich body of evidence suggests that human experts learn to `see' problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}. Novices notice a problem’s `surface features', the literal objects and details made obvious by the task description. Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures', the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.

\begin{figure}
  \includegraphics[width = \textwidth]{figures/text_tsne.png}
  \caption{Left: TSNE embedding of model's `at-a-glance' (before searching for solution) organization of text tasks (neural net activations), after first wake/sleep cycle. Middle: at-a-glance organization after 20 wake/sleep cycles. Right: TSNE visualization of programs actually used to solve tasks; intuitively, how model organizes tasks after searching for a solution. More examples in Fig.~\ref{allTSNE}}\label{textExpertiseCluster}
\end{figure}

\subsection{Programs from visual input}
We consider three domains where the agent must infer a program from an
image (Figure~\ref{visualSpecs}).
First we consider programs that make plans and take actions:  drawing pictures and building towers out of blocks (Sec.~\ref{logoSection}-\ref{towerSection}).

\subsubsection{Programs that draw pictures}\label{logoSection}

Procedural or generative visual concepts 
--- from Bongard problems~\cite{Moscow}, to handwritten characters~\cite{lake2015human,hofstadter1993letter}, to Raven's progressive
matrices~\cite{raven2003raven} --- are studied across AI and cognitive science.  Here we take
inspiration from LOGO Turtle graphics~\cite{turtle}, tasking our agent
with drawing a corpus of 160 images while equipping it with control
over a `pen', along with arithmetic operations on angles and
distances.

Inside its learned library we find interpretable parametric drawing
routines corresponding to the families of visual objects in its
training data, like polygons, circles, and spirals
(Figure~\ref{logoPrimitives}, left) -- without supervision the agent
has learned the basic types of objects in its visual world. It
additionally learns more abstract visual relationships, like
rotational symmetry, which it models by abstracting out a new higher-order
function into its library (Figure~\ref{logoPrimitives}, right).

What does \system dream of?  Prior to learning random programs written using the library
are simple and largely unstructured (Figure~\ref{logoDreams}, left).
After training the samples become richly structured
(Figure~\ref{logoDreams}, right), compositionally recombining latent
building blocks and motifs acquired from the training data. This
offers a visual window into how the generative model bootstraps recognition
model training: as the library grows more finely tuned to the domain, the
neural net gets richer and more highly varied training data.

%% has primitives for moving the pen forward and rotating it (\code{move}),
%% picking the pen up and then putting it down (\code{pen-up}),
%% arithmetic operations on angles and distances,
%% and constants ($2\pi$, 0 through 9, and $\infty$\footnote{To ensure that programs terminate, we set $\infty = 20$}).

\begin{figure}\centering
  \includegraphics[width = 0.3\textwidth]{figures/logo16.png}
  \includegraphics[width = 0.3\textwidth]{figures/tower9.png}
  \includegraphics[width = 0.3\textwidth]{figures/sr16.png}
  \caption{Three domains where the agent infers a program from visual input. \textbf{Left}: 16 (out of 160) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: 9 (out of 112) tower building tasks. Agent writes a program controlling a `hand' that builds the target tower. \textbf{Right}: 16 (out of 200) symbolic regression tasks. Agent writes a program containing continuous real numbers that fits the points along the curve.}\label{visualSpecs}
\end{figure}
\begin{figure}
\centering  \begin{tabular}{cc}
    \toprule
    Parametric drawing routines&Higher-order drawing routine\\\midrule
    \begin{tabular}{rl}
      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
      Circles:
      &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}\\
      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}\\
      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
        Polygons \& Stars:
&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}
    \end{tabular}
    &
    \begin{tabular}{c}
      Rotational Symmetry:\\
          \includegraphics[width = 0.35\textwidth]{figures/rotationalmontage.png}
      \end{tabular}

    \\\bottomrule 
  \end{tabular}
  \caption{Example primitives learned by \system when trained on tasks in Figure~\ref{everyLogo}. Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right). Each row of images on the left is the same code executed with different parameters. Each image on the right is the same code executed with different parameters and with a different subprogram provided as input.}\label{logoPrimitives}
  \end{figure}
\begin{figure}
\centering  \begin{tabular}{c|cc}
  \begin{tabular}{c}
    \includegraphics[width = 0.45\textwidth]{figures/initialDreams/montage.png}
    \end{tabular}&
      \begin{tabular}{c}
        \includegraphics[width = 0.45\textwidth]{figures/finalDreams/cherry_random_montage.png}
        \end{tabular}&
      \rotatebox[origin=c]{90}{Cherry picked\hspace{3cm}Random}
    \end{tabular}
  \caption{Sixteen dreams, or samples, from the DSL before (left) and after (right) training on tasks in Figure~\ref{everyLogo}. Blue: where the agent started drawing. Pink: where the agent ended drawing.}\label{logoDreams}
\end{figure}

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.
The system starts with the same control flow primitives as with
LOGO graphics,
and learns parametric `options' for building blocks towers (Figure~\ref{tower}),
inferring concepts like arches, staircases, bridges and brick walls.
\begin{figure}
\centering  \begin{tabular}{cc}
    \toprule Example learned DSL components&Samples written using final library\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 5cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{stolle2002learning},
    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', built from learned library.}\label{tower}
\end{figure}

\subsubsection{Symbolic Regression}\label{regressionSection}
Here, the agent observes points along the curve of a function, and
must write a program that fits those points.  We initially equip our
learner with addition, multiplication, and division, and task it with
solving 200 symbolic regression problems, each either a polynomial or
rational function.  The recognition model is a convnet that observes
an image of the target function's graph (Fig.~\ref{visualSpecs},
rightmost) --- visually, different kinds of polynomials and rational
functions produce different kinds of graphs, and so the convnet can
look at a graph and predict what kind of function best explains it.  A
key difficulty, however, is that these problems are best solved with
programs containing real numbers.  Our solution to this difficulty is
to enumerate programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.  We
define the likelihood model, $\probability[x|p]$, by assuming a
Gaussian noise model for the input/output examples, and penalize the
use of real-valued parameters using the
BIC~\cite{Bishop:2006:PRM:1162264}.

We learn a DSL containing 13 new functions, mainly templates for
different pieces of polynomials or ratios of polynomials.  The model
also learns to find programs minimizing the number of continuous
parameters --- for example, learning to represent linear functions
with \code{(* real (+ x real))}.  This phenomenon arises from our
Bayesian framing: both the generative model's bias toward shorter
programs, and the likelihood model's BIC penalty.

\subsubsection{Generative Modeling of Text}
Few-shot learning of simple generative concepts comes naturally to
humans, from learning new rules in natural language~\cite{marcus1999rule},
to learning routines for symbols and signs
\cite{lake2015human},
to learning new motor routines for producing words.
We investigate few-shot generative modeling by tasking our
agent with inferring a probabilistic regular expression from a small
number of strings, where these strings are drawn from CSV columns crawled from the web~\cite{LUKE}. Across wake/sleep cycles, our agent learns to learn
regular expressions that describe the structure of a typically occurring text concepts,
like numbers or dates.
Figure~\ref{qualitativeRegex} shows example model outputs on held out generative modeling tasks,
contrasting the full model with ablations that lesion either the library learning or recognition model training.
\begin{figure}
  \includegraphics[width = \textwidth]{figures/regexQualitative.pdf}
  \caption{Qualitative results on probabilistic generative text modeling. Agent observes 5 strings (`Input' column) and infers a probabilistic regex (MAP program--regex primitives highlighted in red), from which it can imagine more examples of the latent text concept. No Library: Dreaming only (ablates library learning). No Rec: Consolidation only (ablates recognition model)}\label{qualitativeRegex}
\end{figure}

@@

\subsection{Quantitative Results on Held-Out Tasks}\label{quantitative}
To evaluate the relative importance of library learning and recognition
model training, we evaluate on held-out testing tasks for each of our
domains, measuring both how many tasks are solved and how long it
takes to solve them across successive wake/sleep iterations
(Fig.~\ref{learningCurves}).  We always solve more held-out tasks --
and generally solve them in less time -- with both components
combined, suggesting a synergy between library learning and
recognition model training. Examining the libraries learned with and without
recognition models reveals a difference in the typical \emph{depth}
of learned subroutines, and this average depth correlates
with the number of solved testing tasks ($r=0.71$, $p < 10^{-8}$).
In contrast, the correlation between the \emph{number} of learned subroutines and number of solved tasks is weaker and not statistically significant ($r=0.20$, $p > 0.05$),
suggesting that the recognition model bootstraps
``better'' libraries, where ``better'' is at least partly explained by the depth
of the learned representation. Figure~\ref{depthMontage} shows typical `deep' libraries learned by
\system and contrasts this depth with \% solved tasks, both with and without the recognition model.
\begin{figure}
  \begin{tabular}{ccc}
    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/text_hits.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/text_time.png}
    \end{tabular}&

@@

\begin{figure}
  \includegraphics[width = \textwidth]{deepMontage.pdf}
  \caption{Top: Learned libraries diagrammed as networks. Arrows point from a learned concept to all other concepts that use it. Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (middle), while total library size is more weakly associated with solving tasks (right). Full model (circles on scatterplots) learns deeper libraries than no-recognition-model ablation (crosses on scatterplots). Each scatter plot point run with a different random seed.}\label{depthMontage}
  \end{figure}

>>>>

\subsection{Conditional programs that manipulate sequences}\label{sequences}
We first apply \system to two classic benchmark domains: list
processing and text editing. In both cases we solve tasks specified by
a conditional mapping (i.e., input/output examples), starting with a
generic functional programming basis, including routines like
\code{map}, \code{fold}, \code{cons}, \code{car}, \code{cdr}, etc.
%% : \code{foldr}, \code{unfold}, \code{if}, \code{map},
%% \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0},
%% \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and
%% \code{is-nil}.

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system typically composes around 20 new library routines, and discovered multiple
higher-order functions. Each round of abstraction built on
concepts discovered in earlier sleep cycles --- for example the
agent first learns the higher-order function \code{filter}, uses
\code{filter} to learn to take the maximum element of a list, then
uses that routine to learn a new component for extracting the
$n^{\text{th}}$ largest element of a list, which it finally uses to
solve a task involving sorting a list of numbers (Figure~\ref{exampleDSL}).
This incremental, modular learning of deep hierarchies of library
components occurs because of the alternation between code writing
(during waking) and code refactoring (during the abstraction phase
of sleep).

@@

\subsection{Generative \& Procedural Programs}
%% We consider three domains where the agent must infer a program from an
%% image (Figure~\ref{visualSpecs}).
%% First we consider programs that make plans and take actions:  drawing pictures and building towers out of blocks (Sec.~\ref{logoSection}-\ref{towerSection}).

\subsubsection{Programs for Smooth Trajectories}\label{regressionSection}
Many procedures that people learn, like as motor routines, can be
thought of as some kind of smooth trajectory. Modeling these
procedures as programs entails jointly inferring their combinatorial
structure and their continuous parameters.  We consider simple
programs generating trajectories described by polynomials (up to
degree 4) and rational functions, and equip our learner with a
primitive for introducing a continuous parameter into the discrete
structure of its programs, which an inner loop optimizes via gradient
descent. We penalize the use of real-valued parameters using a
BIC~\cite{Bishop:2006:PRM:1162264} likelihood model.  We use a convolutional neural
network as a `visual' recognition model, which observes a graph of the
trajectory over time, allowing the agent to `eyeball' the trajectory
prior to writing code that describes it
(Figure~\ref{symbolicRegressionExamples}).  In solving 100 such tasks,
the model learns to find programs minimizing the number of continuous
parameters --- this phenomenon arises from our Bayesian framing: both
the generative model's bias toward shorter programs, and the likelihood
model's BIC penalty.
\begin{figure}
    \includegraphics[width = \textwidth]{figures/sr16_horizontal.png}
  \caption{Sixteen smooth-trajectory tasks. Agent writes a program containing continuous real numbers that fits the points along the curve. Recognition model is a CNN that takes as input a graph of the trajectory over time (shown above). Task is `solved' if program matches points on curve.}\label{symbolicRegressionExamples}
  \end{figure}
\subsubsection{Programs that generate images}\label{logoSection}

Procedural or generative visual concepts 
--- from Bongard problems~\cite{Moscow}, to handwritten characters~\cite{lake2015human,hofstadter1993letter}, to Raven's progressive
matrices~\cite{raven2003raven} --- are studied across AI and cognitive science.  Here we take
inspiration from LOGO Turtle graphics~\cite{turtle}, tasking our agent
with drawing a corpus of 160 images (Figure~\ref{logoMaster}, top) while equipping it with control
over a `pen', along with arithmetic operations on angles and
distances.

Inside its learned library we find interpretable parametric drawing
routines corresponding to the families of visual objects in its
training data, like polygons, circles, and spirals
(Figure~\ref{logoPrimitives}, middle) -- without supervision the agent
has learned the basic types of objects in its visual world. It
additionally learns more abstract visual relationships, like
rotational symmetry, which it models by abstracting out a new higher-order
function into its library.

What does \system dream of?  Prior to learning random programs written using the library
are simple and largely unstructured (Figure~\ref{logoMaster}, bottom left).
After training the samples become richly structured
(Figure~\ref{logoMaster}, bottom right), compositionally recombining latent
building blocks and motifs acquired from the training data. This
offers a visual window into how the generative model bootstraps recognition
model training: as the library grows more finely tuned to the domain, the
neural net gets richer and more highly varied training data.
\begin{figure}
  \centering
  \begin{tabular}{cc}
\toprule     \multicolumn{2}{c}{    Subset of 160 tasks}\\\midrule
    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/horizontalLogoTasks.png}}\\\\

    \midrule  Parametric drawing routines in library&Higher-order drawing routine in library\\\midrule
    \begin{tabular}{rl}
      Semicircle:& \raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_24.png}}\\
      Circles:
      &\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_23.png}}\\
      Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_25.png}}\\
      Greek Spiral:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_10.png}}\\
      S-Curves:&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_11.png}}        \\
        Polygons \& Stars:
&\raisebox{-.5\height}{\includegraphics[width = 0.3\textwidth]{figures/logo_primitives/logo_primitive_12.png}}
    \end{tabular}
    &
    \begin{tabular}{c}
      Radial Symmetry:\\
          \includegraphics[width = 0.35\textwidth]{figures/rotationalmontage.png}
      \end{tabular}

    \\\bottomrule \\\\
    \multicolumn{2}{c}{
      \begin{tabular}{ccc}
        Dreams before learning&Dreams after learning&\\
        \begin{tabular}{c}
          \includegraphics[width = 0.4\textwidth]{figures/initialDreams/montage.png}
        \end{tabular}&
        \begin{tabular}{c}
          \includegraphics[width = 0.4\textwidth]{figures/finalDreams/cherry_random_montage.png}
        \end{tabular}&
        \rotatebox[origin=c]{90}{Cherry picked\hspace{2.5cm}Random}
      \end{tabular}
    }
  \end{tabular}
  \caption{\textbf{Top}: 48 (out of 160) LOGO graphics tasks. Agent writes a program controlling a `pen' that draws the target picture. \textbf{Middle}: Example learned library routines. Agent learns parametric routines for drawing families of curves (left) as well as subroutines that take entire programs as input (right). Each row of images on the left is the same code executed with different parameters. Each image on the right is the same code executed with different parameters and with a different subprogram provided as input. \textbf{Bottom}: Sixteen dreams, or sampled programs, from library before (left) and after (right) learning. Blue: where the agent started drawing. Pink: where the agent ended drawing.}\label{logoMaster}
  \end{figure}

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{tower}, top).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.
The system starts with the same control flow primitives as with
LOGO graphics,
and learns parametric `options' for building blocks towers (Figure~\ref{tower}, bottom left),
inferring concepts like arches, staircases, bridges and brick walls.
\begin{figure}
  \centering  \begin{tabular}{cc}
    \toprule
    \multicolumn{2}{c}{Subset of 112 Tasks}\\\midrule
    \multicolumn{2}{c}{\includegraphics[width = \textwidth]{figures/tower12.png}}\\\midrule \\
    \toprule Example learned DSL components&Samples written using final library\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 6cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 6cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{\textbf{Top}: Twelve (out of 112) tower building tasks. Agent writes a program controlling a `hand' that builds the target tower. \textbf{Bottom left}: Three learned library routines. These components act like parametric options~\cite{stolle2002learning},
    giving higher-level building blocks that the agent can use to plan. \textbf{Bottom Right}: 16 random samples, or `dreams', built from learned library.}\label{tower}
\end{figure}

\subsubsection{Probabilistic Generative Modeling of Text}
Few-shot learning of simple generative concepts comes naturally to
humans, from learning new rules in natural language~\cite{marcus1999rule},
to learning routines for symbols and signs
\cite{lake2015human},
to learning new motor routines for producing words.
We investigate few-shot generative modeling by tasking our
agent with inferring a probabilistic regular expression from a small
number of strings, where these strings are drawn from CSV columns crawled from the web~\cite{LUKE}. Across wake/sleep cycles, our agent learns to learn
regular expressions that describe the structure of a typically occurring text concepts,
like numbers or dates.
Figure~\ref{qualitativeRegex} shows example model outputs on held out generative modeling tasks,
contrasting the full model with ablations that lesion either the library learning or recognition model training.
\begin{figure}
  \includegraphics[width = \textwidth]{figures/regexQualitative.pdf}
  \caption{Qualitative results on probabilistic generative text modeling. Agent observes 5 strings (`Input' column) and infers a probabilistic regex (MAP program--regex primitives highlighted in red), from which it can imagine more examples of the latent text concept. No Library: Dreaming only (ablates library learning). No Rec: Consolidation only (ablates recognition model)}\label{qualitativeRegex}
\end{figure}

@@

\subsection{Analyzing \system across domains}\label{quantitative}
To evaluate the relative importance of library learning and recognition
model training, we evaluate on held-out testing tasks for each of our
domains, measuring both how many tasks are solved and how long it
takes to solve them across successive wake/sleep iterations
(Fig.~\ref{learningCurves}).  We always solve more held-out tasks --
and generally solve them in less time -- with both components
combined, suggesting a synergy between library learning and
recognition model training. Examining the libraries learned with and without
recognition models reveals a difference in the typical \emph{depth}
of learned subroutines, and this average depth correlates
with the number of solved testing tasks ($r=0.71$, $p < 10^{-8}$).
In contrast, the correlation between the \emph{number} of learned subroutines and number of solved tasks is weaker and not statistically significant ($r=0.20$, $p > 0.05$),
suggesting that the recognition model bootstraps
``better'' libraries, where ``better'' is at least partly explained by the depth
of the learned representation. Figure~\ref{depthMontage} shows typical `deep' libraries learned by
\system and contrasts this depth with \% solved tasks, both with and without the recognition model.
\begin{figure}
  \begin{tabular}{ccc}
    \begin{tabular}{c}
      \includegraphics[width = 5cm]{figures/learningCurves/text_hits.png}\\
      \includegraphics[width = 5cm]{figures/learningCurves/text_time.png}
    \end{tabular}&

@@

\begin{figure}
  \includegraphics[width = \textwidth]{deepMontage.pdf}
  \caption{Top: Learned libraries diagrammed as networks. Arrows point from a learned concept to all other concepts that use it. Bottom: Deeper networks solve more testing tasks, both as measured by average depth (left) and maximum depth (middle), while total library size is more weakly associated with solving tasks (right). Full model (stars on scatterplots) learns deeper libraries than no-recognition-model ablation (triangles on scatterplots). Each scatter plot point run with a different random seed.}\label{depthMontage}
\end{figure}

How do \systemEnding's representations of these tasks themselves change, as the model acquires expertise? Within cognitive science, a rich body of evidence suggests that human experts learn to `see' problems differently in their domains -- even before solving a problem, human novices and experts seem to construct fundamentally different initial representations \cite{chi1981categorization}. Novices notice a problem’s `surface features', the literal objects and details made obvious by the task description. Experts, on the other hand, somehow see past these details to pick out each problem’s `deep structures', the underlying principles and concepts that govern its ultimate solution \cite{chi2012seeing,chi2006two}.

\begin{figure}
  \includegraphics[width = \textwidth]{figures/textLogo_tsne.png}
  \caption{Left: TSNE embedding of model's `at-a-glance' (before searching for solution) organization of tasks (neural net activations), after first wake/sleep cycle. Middle: at-a-glance organization after last wake/sleep cycle. Right: TSNE visualization of programs actually used to solve tasks; intuitively, how model organizes tasks after searching for a solution. We convert each program into a feature vector by counting the \# times each library routine is used, and then embed those feature vectors. More examples in Fig.~\ref{allTSNE}}\label{textExpertiseCluster}
\end{figure}
